RDA trait prediction tutorial

### 2023-02-07 choosing-population-genetics.Rmd
caching choosing pop gen info page, update index and yaml files for rda tutorial, fixed few typos rda tutorial

```{r setup, include=FALSE}
#library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(knitcitations)
library(kableExtra)
knitr::opts_chunk$set(fig.width = 10,
               fig.height = 5,
               cache = FALSE)
cite_options(citation_format = "pandoc", max.names = 3, style = "html",
             hyperlink = "to.doc")
```

>>>>

```{r setup, include=FALSE}
#library(knitr)
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(knitcitations)
library(kableExtra)
knitr::opts_chunk$set(fig.width = 10,
               fig.height = 5,
               cache = TRUE)
cite_options(citation_format = "pandoc", max.names = 3, style = "html",
             hyperlink = "to.doc")
```

### 2023-02-07 index.Rmd
caching choosing pop gen info page, update index and yaml files for rda tutorial, fixed few typos rda tutorial

* [Choosing a population genomics sequencing approach](choosing-population-genetics.html)  
* [Whole Genome Resequencing (WGS)](WGS_intro.html)  
  * [Low coverage WGS](https://github.com/therkildsen-lab/genomic-data-analysis/blob/master/lcwgs_data_analysis.md)
* [Reduced Representation Sequencing](RADseq.html) 
* [Poolseq](poolseq.html)

>>>>

* [Choosing a population genomics sequencing approach](choosing-population-genetics.html)  
* [Whole Genome Resequencing (WGS)](WGS_intro.html)  
  * [Low coverage WGS](https://github.com/therkildsen-lab/genomic-data-analysis/blob/master/lcwgs_data_analysis.md)
* [Reduced Representation Sequencing](RADseq.html) 
* [Poolseq](poolseq.html)
* [Redundancy Analysis (RDA) Trait Prediction](rda_tutorial/RDAtraitPredictionTutorial.html)

### 2023-02-07 rda_tutorial/RDAtraitPredictionTutorial.Rmd
caching choosing pop gen info page, update index and yaml files for rda tutorial, fixed few typos rda tutorial

---
title: "RDA trait prediction tutorial"
author: "KE Lotterhos"
date: "2023-01-20"
output:
  pdf_document: default
  html_document: default
---

@@

knitr::opts_knit$set(root.dir = "~/Documents/GitHub/MVP-NonClinalAF/tutorial") 
#change code to your working directory

@@

## RDA trait prediction funciton

@@

```{r function}
rda_trait_pred <- function(rdaobj, env_row, K){
  #rdaobj is RDA object
  #envi row is the row of the environment in the biplot output
  #K is the number of RDA axes
  scores <- scores(rdaobj, choices=1:K)
  ind.sc <- scores$sites
  pred <- matrix(NA, nrow=nrow(ind.sc), ncol=K)
  for (k in 1:K){
    pred[,k] <- ind.sc[,k]*eigenvals(rdaobj)[k]*summary(rdaobj)$biplot[env_row,k]
  }
  trait_pred <- scale(rowSums(pred))
 return(trait_pred) 
}
```

@@

Scale the environmental variables to have a mean of 0 and standard deviation of 1
```{r example}

@@

3. In this case, the first 3 RDA axes explain 80% of the variance. Note that choosing too many axes may result in overfitting. Here is an example of a trait prediction for MAT using the first 3 RDA axes:
```{r}
# Make the trait prediction for MAT (1st row in biplot output)
K = 3 # use 3 RDA axes to make the trait prediciton

@@

Note that the `predict` function and it's variations in the R package `vegan` do not make the same kind of predictions as `rda_trait_pred`. Here are the types of outputs produced by the function `predict` and it's variations:

@@

  # As a side note, it outputs the same thing as the `fitted` funciton
  fitted_values_predict2 <- fitted(rdaout)
  str(fitted_values_predict2)

@@

The `predict` function and it's variations make predictions in RDA space, and therefore do not output the same kind of predictions as `RDA_trait_predict` and Equation 1 in the paper.

@@

heatmap(t(Qr.coeff[,which(muts$causal)]),
        scale="none", col = cm.colors(length(brks)-1), breaks=brks)

>>>>

---
title: "RDA trait prediction tutorial"
author: "KE Lotterhos"
date: "2023-01-20"
---

@@

#change code to your working directory
#knitr::opts_knit$set(root.dir = "~/Documents/GitHub/MVP-NonClinalAF/tutorial") 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)

@@

## RDA trait prediction function

@@

```{r function}
rda_trait_pred <- function(rdaobj, env_row, K){
  #rdaobj is RDA object
  #env_row is the row of the environment in the biplot output
  #K is the number of RDA axes
  scores <- scores(rdaobj, choices=1:K)
  ind.sc <- scores$sites
  pred <- matrix(NA, nrow=nrow(ind.sc), ncol=K)
  for (k in 1:K){
    pred[,k] <- ind.sc[,k]*eigenvals(rdaobj)[k]*summary(rdaobj)$biplot[env_row,k]
  }
  trait_pred <- scale(rowSums(pred))
 return(trait_pred) 
}
```

@@

Scale the environmental variables to have a mean of 0 and standard deviation of 1.
```{r example}

@@

3. In this case, the first 3 RDA axes explain 80% of the variance. Note that choosing too many axes may result in overfitting. Here is an example of a trait prediction for MAT using the first 3 RDA axes:
```{r}
# Make the trait prediction for MAT (1st row in biplot output)
K = 3 # use 3 RDA axes to make the trait prediction

@@

Note that the `predict` function and it's variations in the R package `vegan` do not make the same kind of predictions as `rda_trait_pred`. Here are the types of outputs produced by the function `predict` and its variations:

@@

  # As a side note, it outputs the same thing as the `fitted` function
  fitted_values_predict2 <- fitted(rdaout)
  str(fitted_values_predict2)

@@

The `predict` function and its variations make predictions in RDA space, and therefore do not output the same kind of predictions as `rda_trait_predict` and Equation 1 in the paper.

@@

dev.new()
heatmap(t(Qr.coeff[,which(muts$causal)]),
        scale="none", col = cm.colors(length(brks)-1), breaks=brks)
